---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%",
  fig.width = 5,
  fig.asp = 0.618
)
```

# rosettafish

[![Travis build status](https://travis-ci.org/pbs-assess/rosettafish?branch=master)](https://travis-ci.org/pbs-assess/rosettafish)

rosettafish is an R package to translate fish- and fisheries-related words or short phrases between languages (e.g. French to English, English to French).

This would apply to figure labels to produce separate figures in two languages, or a single figure with labels in two languages. It can also be used for other text (such as table headings), but is not intended for translation of complete sentences.

In particular, it should be useful for automatically producing translated technical figures to go into Research Documents and presentations. This is done when building the figures (in R) and saves someone having to manually edit the figure (in, say, Photoshop). This preserves the quality and the output formats of figures in both languages, and is especially time-saving when the same figure axes are used for multiple figures (e.g. a time series of biomass estimates for multiple model runs of a stock assessment).

The package has a built-in .csv file of English-French translations of common technical fisheries terms (that are often incorrectly translated by generic automatic translators). Users can add to this or use their own file, and are encouraged to add to the ever-expanding list of common terms using the instructions below.

This is currently implemented for French-English translations but could be expanded to other languages (e.g. Inuktitut).

## Installation

You can install rosettafish with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("pbs-assess/rosettafish")
```

## Examples

```{r}
library(rosettafish)
library(ggplot2)
```

```{r}
set.seed(1)
years <- seq(1980, 2018)
df <- data.frame(years, biomass = 2000 * rlnorm(length(years), 0, 1))
```

```{r}
make_plot <- function(french = FALSE) {
  ggplot(df, aes(years, biomass)) + geom_line() +
    xlab(en2fr("year", french)) +
    ylab(en2fr("biomass", french))
}
```

```{r func-figs}
make_plot()
make_plot(french = TRUE)
```

```{r gg-figs}
ggplot(df, aes(years, biomass)) + geom_line() +
  xlab(fr2en("année")) +
  ylab(fr2en("biomasse"))
```

## Instructions for updating the list of translated scientific terms

Fork and clone this repoisitory. Add the English and French terms to the end of the file `inst/extdata/terms.csv`, push to GitHub and submit a pull request. If you do not know what that means, just create an Issue on the GitHub site and give us your new translated terms.

Please note that the 'rosettafish' project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.